Classical probability measures as pullback of quantum probability measures?

### 2022-05-02 pullback.tex
中文摘要.lyx Copied from dissertation.lyx as required by 中興大學人工智慧學程...

\maketitle
It is clear that each step of the measurement process is exactly equal
or just mapped by $\varphi$ except we summarize classical measurement
process as random variable, but summarize quantum measurement process
as observable. Notice that after we fix an orthonormal basis~$\Omega$,
$\varphi$ is actually an invertible map so that given an observable
$\mathbf{O}=\sum_{i=0}^{D-1}\lambda_{i}P_{i}$, we can define an unique
random variable $\sum_{i=0}^{D-1}\lambda_{i}\mathbf{1}_{\varphi^{-1}\left(P_{i}\right)}$
corresponding to $\mathbf{O}$. Therefore, we call this random variable
as the pullback of $\mathbf{O}$ by $\varphi$, and denoted by $\varphi^{*}\mathbf{O}$.\footnote{We might also choose the other way around so that each random variable
has a push forward observable. This is better in a sense that we don't
have the ambiguity that $\varphi^{*}\mathbf{O}$ is might be different
for different orthonormal bases, but every theorem for IVPMs starts
from the quantum part with an observable, so we need the notation
$\varphi^{*}\mathbf{O}$ and fixing an orthonormal basis to disambiguate
$\varphi^{*}\mathbf{O}$ anyway.} According to Gleason's theorem, the table could be extended to simplify
the formula computing expectation values and the the post-measurement-state
postulates after we define a notation to send probability measures
to states and the other way around.

By following the same process, Sec.~\ref{sec:Interval-Uncertainty}
showed the corresponding results for interval-valued probability.
For any QIVPM $\bar{\mu}\colon\events\rightarrow\mathscr{I}$, the
function $\varphi^{*}\bar{\mu}\colon2^{\Omega}\rightarrow\mathscr{I}$
defined by $\left(\varphi^{*}\bar{\mu}\right)\left(E\right)=\bar{\mu}\left(\varphi\left(E\right)\right)$
is a classical IVPM.
\begin{sidewaystable}
\noindent \centering{}\caption{\label{tab:Comparison-interval}Comparison between the classical and
quantum measurement processes with interval-valued probability for
finite dimensional cases}
%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{YYYY}
\toprule 
Name & Classical & Quantum & How does the structure preserve?\tabularnewline
\midrule
\midrule 
Probability measure & $\begin{array}{cccc}
\varphi^{*}\bar{\mu}\colon & 2^{\Omega} & \rightarrow & \mathscr{I}\\
 & E & \mapsto & \bar{\mu}\left(\varphi\left(E\right)\right)
\end{array}$ & $\begin{array}{cccc}
\bar{\mu}\colon & \events & \rightarrow & \mathscr{I}\\
 & E & \mapsto & \bar{\mu}\left(E\right)
\end{array}$ & \tabularnewline
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{sidewaystable}

@@

Given a finite sample space~$\Omega$ representing all possible outcomes
of a process, and its power set $2^{\Omega}$ as the classical event
space, a classical probability measure~$\mu$ maps every event $E\subseteq\Omega$
to a number $\mu\left(E\right)\in\left[0,1\right]$ specifying how
likely one of the outcomes in $E$ will happen. To maintain the coherence,
$\mu$ is subject to the following constraints: $\mu\left(\emptyset\right)=0$,
$\mu\left(\Omega\right)=1$, and $\mu\left(\overline{E}\right)=1-\mu\left(E\right)$,
where $\overline{E}$ is the complement of $E$.  Moreover, we require
$\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$
for each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.

@@

Since we may rotate the inhomogeneous magnetic field any degree~$\theta$,
it is nature to represent the two bent outcomes as $\sin\theta\ket{0}+\cos\theta\ket{1}$
and $\cos\theta\ket{0}-\sin\theta\ket{1}$. Then, $\ket{\ps}$ and
$\ket{\ms}$ can be represented as $\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}$
and $\ket{\ms}=\frac{\ket{0}-\ket{1}}{\sqrt{2}}$. Conventionally
\cite{peres1995quantum}, we will then postulate any normalized complex
vector spanned by $\ket{0}$, $\ket{1}$, and $\ket{2}$ could be
a measurement outcome, any orthonormal basis form a sample space,
and sending a particle to this orthonormal basis inducing a classical
probability measure. In general, as long as sending the same particle,
the probability of the same event in different experiments should
always be the same, i.e., giving two induced classical probability
measures $\mu'\colon2^{\Omega'}\rightarrow\left[0,1\right]$ and $\mu''\colon2^{\Omega''}\rightarrow\left[0,1\right]$,
the probabilities of getting $E\subseteq\Omega'\cap\Omega''$ are
the same
\begin{equation}
\mu'\left(E\right)=\mu''\left(E\right)\,.\label{eq:mu-equal-same-event}
\end{equation}
\footnote{This fact is equivalent to the fact that commuting observables could
be measured simultaneously which is essential to define contextuality
and will be explained in Sec.~\ref{sec:Kochen-Specker}... Or maybe
there is a gap between outcomes in a few types of experiments and
every normalized vector in a Hilbert space could be an outcome, but
inducing latter from axioms seems a bit too much here\ldots}

>>>>

\maketitle

@@

Given a finite sample space~$\Omega$ representing all possible outcomes
of a process, and its power set $2^{\Omega}$ as the classical event
space, a classical probability measure~$\mu$ maps every event $E\subseteq\Omega$
to a number $\mu\left(E\right)\in\left[0,1\right]$ specifying how
likely one of the outcomes in $E$ will happen. To maintain the coherence,
$\mu$ is subject to the following constraints: $\mu\left(\emptyset\right)=0$,
$\mu\left(\Omega\right)=1$, and $\mu\left(\Omega\backslash E\right)=\mu\left(\overline{E}\right)=1-\mu\left(E\right)$,
where $\Omega\backslash E$ is the complement of $E$, and we may
simply write $\overline{E}$ if the sample space~$\Omega$ is clear
from the context.  Moreover, we require $\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$
for each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.

@@

Since we may rotate the inhomogeneous magnetic field any degree~$\theta$,
it is nature to represent the two bent outcomes as $\sin\theta\ket{0}+\cos\theta\ket{1}$
and $\cos\theta\ket{0}-\sin\theta\ket{1}$. Then, $\ket{\ps}$ and
$\ket{\ms}$ can be represented as $\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}$
and $\ket{\ms}=\frac{\ket{0}-\ket{1}}{\sqrt{2}}$. Conventionally
\cite{peres1995quantum,schwinger2001quantum}, we will then postulate
any normalized complex vector spanned by $\ket{0}$, $\ket{1}$, and
$\ket{2}$ could be a measurement outcome, any orthonormal basis form
a sample space, and sending a particle to this orthonormal basis inducing
a classical probability measure. In general, as long as sending the
same particle, the probability of the same event in different experiments
should always be the same, i.e., giving two induced classical probability
measures $\mu'\colon2^{\Omega'}\rightarrow\left[0,1\right]$ and $\mu''\colon2^{\Omega''}\rightarrow\left[0,1\right]$,
the probabilities of getting $E\subseteq\Omega'\cap\Omega''$ are
the same
\begin{equation}
\mu'\left(E\right)=\mu''\left(E\right)\,.\label{eq:mu-equal-same-event}
\end{equation}
\footnote{This fact is equivalent to the fact that commuting observables could
be measured simultaneously which is essential to define contextuality
and will be explained in Sec.~\ref{sec:Kochen-Specker}... Or maybe
there is a gap between outcomes in a few types of experiments and
every normalized vector in a Hilbert space could be an outcome, but
inducing latter from axioms seems a bit too much here\ldots}
